---
title: "Analysis of Hospital Data - Sean Huang"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(dplyr)
library(lmtest)
library(car)


hospital <- read.csv("hospital_data.csv")
hospital

```

# Explanatory Data Analysis

```{r}
plot(hospital$Beds, hospital$InfctRsk, ylab = "InfctRsk", xlab = "Beds")
```

```{r}
plot(hospital$Facilities, hospital$InfctRsk, ylab = "InfctRsk", xlab = "Facilities")
```

```{r}
plot(hospital$MedSchool, hospital$InfctRsk, ylab = "InfctRsk", xlab = "MedSchool")
```

```{r}
pairs(~ Beds + Facilities + MedSchool, data=hospital)
```
There seems to be trend between beds to infection risk and facilities to infection risk, most likely it is quadratic since there is a bit of curve. Meanwhile it is a bit difficult to determine a relationship for Med School to infection risk since it is a categorical variable. In addition, there might be a problem of multicollinearity between beds and facilities.


# Methodology

## Model 1 

```{r}
model1 <- lm(hospital$InfctRsk ~ hospital$Beds + hospital$Facilities + hospital$MedSchool)
summary(model1)
vif(model1)
```
The first model ran with all variables and with no higher order. There appears to have multicollinearity between beds and facilities so will rerun without either variable.

## Model 2 - Beds and MedSchool Only

```{r}
model2 <- lm(hospital$InfctRsk ~ hospital$Beds + hospital$MedSchool)
summary(model2)
vif(model2)
```

## Model 3 - Facilities and MedSchool Only

```{r}
model3 <- lm(hospital$InfctRsk ~ hospital$Facilities + hospital$MedSchool)
summary(model3)
vif(model3)
```
Since the third model had the higher adjusted $R^2$ value or explanatory power, the third model is chosen.

## Checking the assumptions for Model 3

### Normality
```{r}
plot(model3, which=2)
shapiro.test(residuals(model3))
```
Normality is satisfied.

### Constant variance
```{r}
plot(model3$fitted.values, model3$residuals, xlab='Residuals', ylab='Fitted Value')
abline(h=0,lty=2)
bptest(model3)
```
Constant variance is satisfied.

### Independence
```{r}
durbinWatsonTest(model3)
```
Independence is satisfied.

### Mean of error
```{r}
crPlots(model3)
```
There seems to be a problem with facilities as it does not follow a linear trend instead a quadratic term might be needed.


## Model 4 - Facilities^2

```{r}
model4 <- lm(hospital$InfctRsk ~ hospital$Facilities + I(hospital$Facilities^2) + hospital$MedSchool)
summary(model4)
```

The fourth model seems the best so far with the addition of the quadratic term.


## Model 5 - Interaction

```{r}
model5 <- lm(hospital$InfctRsk ~ hospital$Facilities * hospital$MedSchool)
summary(model5)
```
There does not seem to be an interaction between Facilities and MedSchool. 

## Model 6 - Interaction with Higher Order Term

```{r}
model6 <- lm(hospital$InfctRsk ~ hospital$Facilities * I(hospital$Facilities^2) * hospital$MedSchool)
summary(model6)
```
There does not seem to be an interaction between Facilities and MedSchool.

## Model 7 - Removed MedSchool since it is insignificant

```{r}
model7 <- lm(hospital$InfctRsk ~ hospital$Facilities + I(hospital$Facilities^2))
summary(model7)
```

## Final Model

```{r}
plot(hospital$Facilities, hospital$InfctRsk, ylab = "InfctRsk", xlab = "Facilities")
lines(sort(hospital$Facilities), fitted(model7)[order(hospital$Facilities)], col='red')
```

The final model chosen is:
$InfctRsk = 0.748564 + 0.142225*Facilities - 0.001210*Facilities^2$

The infection risk when there is no facility (or facility = 0) is 0.748564.
For each increase of a facility, infection risk is both increased by 0.142225 and decreased by $0.001210*Facilities^2$. For example, if facility is 1, then infection risk is increased by 0.120445.

### Outliers and Influencial Results

```{r}
plot(seq(1,113), rstandard(model7), ylab = "Standardized Residuals",
xlab = "Observation ID")
plot(cooks.distance(model7), ylab = "Cook's Distance",
xlab = "Observation ID")
```
While there appears to be outliers, they are of not much concern since Cook's Distance shows that they are not influential.

## Appendix

```{r, include=TRUE}
library(tidyverse)
library(dplyr)
library(lmtest)
library(car)

hospital <- read.csv("hospital_data.csv")
hospital

# Explanatory Data Analysis
## Scatter Plots
plot(hospital$Beds, hospital$InfctRsk, ylab = "InfctRsk", xlab = "Beds")
plot(hospital$Facilities, hospital$InfctRsk, ylab = "InfctRsk", xlab = "Facilities")
plot(hospital$MedSchool, hospital$InfctRsk, ylab = "InfctRsk", xlab = "MedSchool")
## Correlation Plots
pairs(~ Beds + Facilities + MedSchool, data=hospital)

# Methodology
## Model 1 
model1 <- lm(hospital$InfctRsk ~ hospital$Beds + hospital$Facilities + hospital$MedSchool)
summary(model1)
vif(model1)
## Model 2 - Beds and MedSchool Only
model2 <- lm(hospital$InfctRsk ~ hospital$Beds + hospital$MedSchool)
summary(model2)
vif(model2)
## Model 3 - Facilities and MedSchool Only
model3 <- lm(hospital$InfctRsk ~ hospital$Facilities + hospital$MedSchool)
summary(model3)
vif(model3)
### Checking the assumptions for Model 3
#### Normality - QQ Plot and Shapiro-Wilkes
plot(model3, which=2)
shapiro.test(residuals(model3))
#### Constant variance - Residuals vs Fitted Plot and Breush-Pagan
plot(model3$fitted.values, model3$residuals, xlab='Residuals', ylab='Fitted Value')
abline(h=0,lty=2)
bptest(model3)
#### Independence - Durbin Watson
durbinWatsonTest(model3)
#### Mean of error
crPlots(model3)
## Model 4 - Facilities^2
model4 <- lm(hospital$InfctRsk ~ hospital$Facilities + I(hospital$Facilities^2) + hospital$MedSchool)
summary(model4)
## Model 5 - Interaction
model5 <- lm(hospital$InfctRsk ~ hospital$Facilities * hospital$MedSchool)
summary(model5)
## Model 6 - Interaction with Higher Order Term
model6 <- lm(hospital$InfctRsk ~ hospital$Facilities * I(hospital$Facilities^2) * hospital$MedSchool)
summary(model6)
## Model 7 - Removed MedSchool since it is insignificant
model7 <- lm(hospital$InfctRsk ~ hospital$Facilities + I(hospital$Facilities^2))
summary(model7)

# Final Model Plot
plot(hospital$Facilities, hospital$InfctRsk, ylab = "InfctRsk", xlab = "Facilities")
lines(sort(hospital$Facilities), fitted(model7)[order(hospital$Facilities)], col='red')

# Outliers
plot(seq(1,113), rstandard(model7), ylab = "Standardized Residuals",
xlab = "Observation ID")
# Cook's Distance
plot(cooks.distance(model7), ylab = "Cook's Distance",
xlab = "Observation ID")
```